### 2023-09-06 article-pl.tex
Initial version of article

% !TEX encoding = UTF-8 Unicode

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a proceedings volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[conference]{IEEEtran}
%
% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%

% to typeset URLs, URIs, and DOIs
% https://www.overleaf.com/learn/latex/Hyperlinks
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Overleaf Example},
    pdfpagemode=FullScreen,
    }

\def\UrlFont{\rmfamily}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[textsize=scriptsize]{todonotes}
\usepackage{gensymb}
\usepackage{array,multirow}
\usepackage{caption} 
\captionsetup[table]{skip=10pt}
\usepackage{makecell}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{bm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\usepackage{graphicx}
\graphicspath{ {./img/} }

\renewcommand{\labelenumii}{\theenumii}
\renewcommand{\theenumii}{\theenumi.\arabic{enumii}.}

\begin{document}
%
\title{Duckietown linie folower PID and AI}

\author{
\IEEEauthorblockN{1\textsuperscript{st} Marek Długosz}
\IEEEauthorblockA{\textit{AGH University of Science and Technology}\\
Cracow, Poland \\
mdlugosz@agh.edu.pl \\
0000-0001-6827-9149}\\[0.4cm]
\IEEEauthorblockN{3\textsuperscript{rd} Marcin Szelest}
\IEEEauthorblockA{\textit{IEEE Senior Member}}
\IEEEauthorblockA{\textit{AGH University of Science and Technology} \\
Cracow, Poland \\
mszelest@agh.edu.pl\\
0000-0002-0522-1270}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Paweł Skruch}
\IEEEauthorblockA{\textit{IEEE Senior Member}}
\IEEEauthorblockA{\textit{AGH University of Science and Technology} \\
Cracow, Poland \\
pawel.skruch@agh.edu.pl\\
0000-0002-8290-8375}\\
\IEEEauthorblockN{4\textsuperscript{th} Artur Morys-Magiera}
\IEEEauthorblockA{\textit{AGH University of Science and Technology} \\
Cracow, Poland \\
amorys@student.agh.edu.pl\\
0000-0002-2137-8841}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}

\end{abstract}

% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\begin{IEEEkeywords}
\end{IEEEkeywords}
%
\section{Introduction}

\section{Conclusion}\label{sec:conclusion}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}