### 2023-09-20 article-en.tex
English version correction

% !TEX encoding = UTF-8 Unicode

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a proceedings volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[conference]{IEEEtran}
%
% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%

% to typeset URLs, URIs, and DOIs
% https://www.overleaf.com/learn/latex/Hyperlinks
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
    pdftitle={Application of PID controller and CNN to control Duckiebot robot},
    pdfpagemode=FullScreen,
    }

\def\UrlFont{\rmfamily}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[textsize=scriptsize]{todonotes}
\usepackage{gensymb}
\usepackage{array,multirow}
\usepackage{caption} 
\usepackage{subcaption}
\captionsetup[table]{skip=10pt}
\usepackage{makecell}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{bm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\usepackage{graphicx}
\graphicspath{ {./img/} }

\renewcommand{\labelenumii}{\theenumii}
\renewcommand{\theenumii}{\theenumi.\arabic{enumii}.}

\begin{document}
%
\title{Application of PID controller and CNN to control Duckiebot robot}
% \title{Application of PID and CNN controllers to control the Duckiebot robot}

\author{
\IEEEauthorblockN{1\textsuperscript{st} Marek Długosz}
\IEEEauthorblockA{\textit{AGH University of Science and Technology}\\
Cracow, Poland \\
mdlugosz@agh.edu.pl \\
0000-0001-6827-9149}\\[0.4cm]
\IEEEauthorblockN{3\textsuperscript{rd} Marcin Szelest}
\IEEEauthorblockA{\textit{IEEE Senior Member}}
\IEEEauthorblockA{\textit{AGH University of Science and Technology} \\
Cracow, Poland \\
mszelest@agh.edu.pl\\
0000-0002-0522-1270}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Paweł Skruch}
\IEEEauthorblockA{\textit{IEEE Senior Member}}
\IEEEauthorblockA{\textit{AGH University of Science and Technology} \\
Cracow, Poland \\
pawel.skruch@agh.edu.pl\\
0000-0002-8290-8375}\\
\IEEEauthorblockN{4\textsuperscript{th} Artur Morys-Magiera}
\IEEEauthorblockA{\textit{AGH University of Science and Technology} \\
Cracow, Poland \\
amorys@student.agh.edu.pl\\
0000-0002-2137-8841}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
The paper presents the design and practical implementation by students of a control system
using a classic PID controller and a controller using artificial neural networks.
The control object is a Duckiebot robot, and the task it is to perform is to drive the robot along a designated line (\emph{line follower}). The purpose of the proposed activities is to familiarize students with the advantages and disadvantages of the two controllers used and for them to acquire the ability to implement control systems in practice. 
The article briefly describes how the two controllers work, how to practically implement them, and how to practically implement the exercise.
\end{abstract}

% We would like to encourage you to list your keywords within
% the abstract section using the \keywords{...} command.
\begin{IEEEkeywords}
autonomouse robot, PID controller, convolutional neural network, line follower
\end{IEEEkeywords}
%
\section{Introduction}
Recent years have seen a renaissance of sorts in the use of artificial neural networks. Various types of network architectures have been created and made generally available, as well as their training methods have been improved. All this has led to the application of artificial neural networks being on a rise, particularly successfully, in automatic control systems \cite{algor2030973, 786109}. The use of neural networks in automation systems is not their typical application because very often control can be implemented using well-known and described classical control algorithms (PID, LQR, MPC, etc.). This provides an opportunity to compare the performance of two different technologies and find their advantages and disadvantages. In the process of educating students, the design and implementation by them of a control system using classical control algorithms and artificial neural networks is very interesting from both a scientific and didactic point of view.

An important element of student education that increases the attractiveness of the classes themselves, providing the opportunity to solve practical problems, is the use of real robots. One project that provides a complete robotics platform (hardware, software, simulator, and teaching materials) is the Duckietown project.
The Duckietown project was created in 2016 at the Massachusetts Institute of Technology (MIT) as part of a class for students \cite{paull2017duckietown}. It is an open project that aims to promote education and research related to robot autonomy, especially in the context of autonomous cars. As part of the project, the authors designed entities like Duckiebot robots, mock-ups of the Duckietown cities the robots need to navigate through, and additional infrastructure, such as the Watchtowers, to support the navigation process.
The project also includes the development of simulation software called Duckiebot-GYM, which allows the simulation of robots in a virtual environment \cite{gym_duckietown}. The most recent, second generation of Duckiebot robots, is using the Jetson Nano single-board computer for control. Using the Jetson Nano platform opens the door to using advanced neural networks with dedicated GPU support to control the Duckiebot robot.

The article is organized as follows: Section \ref{sec:problem-definition} defines the problem to be solved, Section \ref{sec:student-activity} describes the student activities that need to be carried out to solve the problem. Section \ref{sec:robot-duckiebot} contains a brief design of Duckietown, the structure and control of the Duckiebot robot, sections \ref{sec:pid-controller} and \ref{sec:nn-controller} describe how to solve the problem using a PID controller and a controller using artificial neural networks. The \ref{sec:practical-info} section gives some practical tips to help with the task, and finally the \ref{sec:conclusion} section provides conclusions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Problem definition}\label{sec:problem-definition}
As part of the didactic process, students have the goal of designing an automatic control system for a differentially driven robot that performs the functionality of moving along a designated route. The route of the robot is marked with a colored line (e.g. red, yellow, etc.) distinguishable from the ground (usually black), on which the robot moves. The shape of the trajectory is a closed contour. The control system should generate a control signal to minimize the total deviation error of the robot from the preset trajectory. This type of task is a typical scenario implemented in various types of autonomous robots and is often referred to as the \emph{line follower}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Student activities}\label{sec:student-activity}
Within the problem defined in the \ref{sec:problem-definition} section, students will have to demonstrate adequate theoretical preparation and practical skills such as: knowledge of the principle of operation of the PID controller, knowledge of the principle of operation of artificial neural networks, knowledge of operations related to image transformations, ability to implement the discrete version of the PID controller, skills to implement selected structures of neural networks, ability to implement operations of image transformations, knowledge of issues related to the selection of parameters of the PID controller, knowledge of issues related to training neural networks and preparing data for their training, evaluation of the correctness of the training process of the network. A seemingly simple algorithm to control an uncomplicated robot that realizes the functionality of \emph{line follower} generates a number of tasks, performance of which requires adequate theoretical and practical preparation from students. It should also be noted that the problems that students have to solve during the project are typical problems that also occur in the programming of other types of autonomous robots.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Duckiebot robot}\label{sec:robot-duckiebot}
The Duckiebot robot is a small two-wheeled robot that is capable of moving on flat surfaces. Two modules containing an electric motor, a gearbox and an encoder were used to drive the robot. The Duckiebot robot is equipped with the following sensors: an RGB camera, a distance sensor and an IMU inertial sensor \cite{gupta2022low}. A Jetson NANO single-board computer is used to control the robot. In addition, the Duckiebot robot is equipped with four RGB LEDs which can be used, for instance, to signal the state the robot is currently in. The Duckiebot robot is designed for autonomous control and can move without human intervention, based on camera data and appropriate control algorithms. The figure \ref{fig:duckiebot-3d} shows the Duckiebot robot in three projections.

\begin{figure}[h]
    \centering
    \includegraphics[width=.9\columnwidth]{duckiebot-blue-3d}
    \caption{Duckiebot robot (\url{http://duckietown.org}).}
    \label{fig:duckiebot-3d}
\end{figure}

A consequence of using two independent electric motors in the design of the Duckiebot robot is the way in which the direction of the robot's movement is controlled.
The Duckiebot robot makes turns by changing the speeds of the electric motors depending on which direction it wants to turn and how large the turning radius should be.
Two signals are used to control the robot: $v$-value of the linear speed and $omega$-value of the rotational speed along the robot's vertical Z axis, see Fig. \ref{fig:duckiebot-kinematics}.

\begin{figure}[h]
    \centering
    \includegraphics[width=.8\columnwidth]{dt-kinematic}
    \caption{Duckiebot robot kinematic system.}
    \label{fig:duckiebot-kinematics}
\end{figure}

Given the values of $(v, \omega)$ (determined, for example, by the controller), we can calculate the rotational speeds of the robot's wheels from the formulas \eqref{eq:omega-L} and \eqref{eq:omega-R}:

\begin{eqnarray}
	\omega_L & = & \cfrac{v-L\omega}{R} \label{eq:omega-L} \\
	\omega_R & = & \cfrac{v+L\omega}{R} \label{eq:omega-R}
\end{eqnarray}
\\
This type of Duckiebot robot control system is very commonly used in robotics and is called \emph{differential drive} \cite{siegwart2011introduction}. The advantage of such a kinematic system for the motion of the Duckiebot robot is the simplicity of the design (no torsion axis mechanisms of the wheels), while the disadvantage is the need to constantly control the rotational speed of the robot's wheels ($\omega_R$, $\omega_L$) so that it moves along the specified route.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Classic implementation with a PID controller}\label{sec:pid-controller}
The classic Duckiebot robot control system uses a PID controller. The task of the control system is to generate such a control signal $u_i$ to minimize the value of the error signal $e_i$, where $i$ corresponds to consecutive time moments. Figure \ref{fig:pid-pipeline} shows a block diagram of the Duckiebot robot control system that uses the PID controller and the robot camera as a sensor.

\begin{figure}[h]
    \centering
    \includegraphics[width=.8\columnwidth]{PipeLinePID}
    \caption{Schematic of Duckiebot robot control system using PID controller.}
    \label{fig:pid-pipeline}
\end{figure}

The error signal $e_i$ is defined as the difference between the center of the horizontal axis of the camera image and the point in the analyzed image region contained within the line along which the Duckiebot robot should move; see Fig. \ref{fig:error-definition}.

\begin{figure}[h]
    \centering
    \includegraphics[width=.95\columnwidth]{ErrorDefinition.png}
    \caption{Error determination for the control system implementing the \emph{line follower} functionality for the Duckiebot robot.}
    \label{fig:error-definition}
\end{figure}

The control system of the Duckiebot robot is a discrete system (implemented on Jetson NANO computer), so the discrete version of the PID controller has to be used. Taking into account the drive type of the Duckiebot robot \emph{differential drive robot}, the control signal $u_i$ generated by the PID controller is equivalent to the rotational speed $u_i$ along the Z axis of the robot.

\subsection{PID controller}
One of the best known and most widely used controllers in automatic control systems is the PID controller. This is primarily due to its versatility, ease of implementation in both analog and digital form, and the relatively simple and intuitive way of tuning the PID controller parameters. The discrete version of the PID controller that was used in the Duckietown robot control system is given by the equation \eqref{eq:pid}:

\begin{equation}
u_i = K_p e_i + K_i \sum_{i=0}^k{e_i \Delta t} + K_d \cfrac{e_i-e_{i-1}}{\Delta t}
\label{eq:pid}
\end{equation}
%https://github.com/duckietown/mooc-exercises/blob/c476a67fea9836beab99eed23cfc1a83c77af7bb/modcon/solution/05-PID-Control/PID_controller.ipynb
where: $K_p$ amplification of the proportional part, $K_i$ amplification of the integrating part, $K_d$ amplification of the differentiating part, $e_i$ error at the $i$-th instant of time, $\Delta t$ time step \cite{aastrom2021feedback}.

\subsection{Error measurement}
The value of the error signal $e_i$ is determined from the images transmitted by the Duckiebot robot camera. To determine the value of the $e_i$ signal, it is necessary to perform several operations on the transmitted images (see Figure \ref{fig:pid-pipeline}). 
The first operation is to convert the image from the RGB color space to the HSV space (\emph{Hue}, \emph{Saturation}, \emph{Value}). This conversion of between color spaces of images is mainly aimed at better color discrimination and recognition.
Another operation is to find a line of certain color in the image, along which the Duckiebot robot should move. Such an operation is to leave only the line marking the robot's route in the image and remove all other elements (e.g. background) from the image. 
The next operation is to reduce the area being analyzed by performing a cropping operation. This operation aims to find the exact center of the line along which the Duckiebot robot will move. 
The center of the line is found as the coordinates of the center of gravity obtained, as a result of the previous transformations, of the line that determines the route of the Duckiebot robot. Figure \ref{fig:img-transformation-summary} shows the result of the specific transformations performed on the robot's camera image to determine the value of the position error $e_i$.

\begin{figure}[hbt!]
    \centering
    \includegraphics[width=.95\columnwidth]{img-transformation.png}
    \caption{Operations performed on robot camera images to determine position error  $e_i$.}
    \label{fig:img-transformation-summary}
\end{figure}

Finally, the position error value $e_i$ is normalized to the interval $e_i \in [-1;1]$, so as to make the result independent of the resolution (width, height) of the images captured from the camera.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Implementation using neural networks}\label{sec:nn-controller}
The second method of generating a control signal $u_i$ uses a convolutional neural network (\emph{convolutional neural network} CNN). The block diagram of the control system using neural networks is shown in Fig. \ref{fig:cnn-pipeline}.

\begin{figure}[h]
    \centering
    \includegraphics[width=.8\columnwidth]{NNPipeline3}
    \caption{Schematic of Duckiebot robot control system using convolutional neural networks.}
    \label{fig:cnn-pipeline}
\end{figure}

Convolutional neural networks are very effective in the tasks of pattern recognition and image processing. CNNs repeatedly use the convolution operation, which allows for extracting characteristic features (edges, textures, shapes) from input images on the basis of which a classical neural network can judge on "what is in the image". Convolutional neural networks are used in various fields, such as image recognition, text analysis, speech processing, facial recognition, control of vehicles, diagnosing diseases from medical images, analyzing satellite images, etc. \cite{li2021survey}, \cite{rawat2017deep}, \cite{almasi2020robust}.

When a CNN is used to control a Duckiebot robot, such a network receives images from the camera as input (after performing several transformations) based on which it generates a $u_i$ control signal which is interpreted as the $\omega_i$ speed along the robot's Z-axis; see Fig. \ref{fig:cnn-pipeline}.

\subsection{Convolutional Neural Network}
The proposed structure of the convolutional neural network is shown in Fig. \ref{fig:cnn-structure} and it should be noted that this is one of the possible network structures that can be used.

\begin{figure}[h]
    \centering
    \includegraphics[width=.5\columnwidth]{cnn2}
    \caption{Structure of the convolutional neural network used to control the Duckiebot robot.}
    \label{fig:cnn-structure}
\end{figure}

The network consists of the first convolutional layer, the next layer is the \emph{MaxPooling} layer. Its purpose is to reduce the size of the input data. The next layers of the network are the convolution layer and the \emph{MaxPooling} layer, and this arrangement of layers is repeated twice. Each time, the convolution layers extract further features of the input image relevant to the control of the Duckiebot robot, while the \emph{MaxPooling} layers reduce the size of the feature maps. The function of the next layer \emph{Dropout} is randomly turning off individual neurons of the network, to prevent the phenomenon of \emph{overfitting} and to further support the generalization ability of the neural network. Before the data is sent to the input of the actual neural network, it must be flattened, which is the reasoning behind placing a \emph{Flatten} layer: to transform multidimensional data into a one-dimensional input vector to pass to the next, \emph{Dense} layer. The \emph{Dense} layer is responsible for performing abstract data representation tasks, e.g. may be used for classification, regression. The output of the CNN is the value of the signal $\omega$ in the interval $[-8 ; -8]$, therefore, the activation function from the \emph{Dense} output layer is the \emph{tanh} function.

\subsection{Data preparation}
A neural network is only as good as the data with which it was trained. Therefore, a very important step in working with neural networks is to prepare a good set of training data. In case of the Duckiebot robot and the task it is supposed to perform (\emph{line follower}), the training data for the network consist of pairs: the image from the camera and the corresponding correct $\omega$ value, see Fig. \ref{fig:cnn-pipeline}. The data necessary to train the network can be obtained by controlling the Duckiebot robot manually (e.g., with a gamepad) or by using a previously developed control system with a PID controller by adding a script that writes the data into a log file.

In addition to the appropriate amount of logged data for training the neural network, it is important that the data is sufficiently diverse, that is, that it covers the largest possible number of cases, situations, events in which the Duckiebot robot may find itself. Figure \ref{fig:cnn-data-hist1} shows a histogram of an example dataset that has been logged.

\begin{figure}[h]
	\begin{subfigure}{.475\columnwidth}
    	\centering
    	\includegraphics[width=1\columnwidth]{h1}
    	\caption{}
    	\label{fig:cnn-data-hist1}
    \end{subfigure}
    \begin{subfigure}{.475\columnwidth}
    	\centering
    	\includegraphics[width=1\columnwidth]{h2}
    	\caption{}
    	\label{fig:cnn-data-hist2}
    \end{subfigure}

    \caption{Histogram of data recorded before analysis (a) and after analysis (b).}
    \label{fig:cnn-data-hist}
\end{figure}

It can be clearly seen that the dominant value in this data set for the variable $\omega$ is close to $0$, and if you use such data to train a CNN that is supposed to control the Duckiebot robot, it is likely that the robot would drive straight very well, while it would have problems turning. In such a case, after analyzing the data, samples with values that significantly dominate other values should be discarded from the training set, obtaining a more representative training data set, see Fig. \ref{fig:cnn-data-hist2}.

Similarly to control of the Duckiebot robot using a PID controller, it is necessary to adequately prepare the camera images when controlling the Duckiebot robot using neural networks. The various stages of processing Duckiebot robot camera images are shown in Figure \ref{fig:cnn-pipeline} (yellow blocks) and are: cropping the image to the area of interest, conversion to YUV color space (Y-luminance, UV color, chrominance), filtering using the Gauss filter, resizing the image and normalizing the values of individual image channels to the interval $[0;1]$. The result of the various operations can be seen in Figure \ref{fig:cnn-image-prepare}.

\begin{figure}[h]
    \centering
    \includegraphics[width=.95\columnwidth]{nn_prepare_image}
    \caption{Operations performed on images from the camera of the Duckiebot robot controlled before the application of the CNN controller.}
    \label{fig:cnn-image-prepare}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Practical advice}\label{sec:practical-info}
The Duckietown robot software was implemented using the Robot Opertion System framework (\emph{ROS}) \cite{quigley2015programming}. It is currently one of the most popular frameworks for programming robots of various kinds. The consequence of choosing the ROS framework is the choice of programming language, and in the case of Duckiebot robots it is Python. The Duckietown project also defines how to develop custom software for robots and enforces the use of containerization technology. This is a very good practice in accordance with the current software development trends. A detailed description together with examples of creating software for Duckie robots is very well described in the project documentation \cite{Daniele:ff}. 
The use of the ROS framework also makes it possible to efficiently create and test software without having to constantly upload new versions to the Duckiebot robot. By configuring the host computer appropriately so that it connects to the ROS server running on the selected Duckiebot robot, you can implement and test your own software conveniently and efficiently, see Fig. \ref{fig:env-develop}.

\begin{figure}[h]
    \centering
    \includegraphics[width=.8\columnwidth]{controll_schema2}
    \caption{Setting up a development environment for the Duckiebot robot.}
    \label{fig:env-develop}
\end{figure}

The proposed controllers should be implemented as ROS nodes. Nodes should read images from the robot's camera (sent as ROS messages) and generate a $\omega$ control signal as output. To simplify the control of the Duckiebot robot, it can be assumed that it moves at a constant linear speed $v=const$.

Convolutional neural networks can be created, trained, and tested using the \emph{keras} and \emph{tensorflow} libraries in the Python programming language. The advantage of the \emph{keras} and \emph{tensorflow} modules is also a very good documentation with a very large number of examples, as well as a large community of developers gathered around them.

It is also necessary to pay attention to the correct preparation of data for training neural networks, each camera image must correspond to a single, correct value of the $\omega$ signal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusions}\label{sec:conclusion}
The PID controller is used successfully in many control systems, and its implementation is relatively simple. There are also a number of methods and algorithms for adjusting controller parameters for this type of controller. PID controllers, on the other hand, are not free of disadvantages. One of them is the requirement of prior knowledge of, even roughly, the model of the process one wants to control. Thus, it is necessary to identify both the structure of the process model and its parameters. Identification tasks are complex tasks, requiring a great deal of knowledge about the nature of the process itself. There are also methods for identifying process models based on the results of practical experiments, however sometimes it may not be possible to conduct such experiments. When using a PID controller, one should also be aware that it was developed for processes, operation of which can be described by linear models. Unfortunately, the behavior of the vast majority of dynamic systems is described by non-linear models. The consequence of this fact is that, in such cases, the PID controller works using linear approximations of nonlinear systems, which can lead to various errors, inaccuracies, etc.

Unlike the classic PID controller, controllers using artificial neural networks do not need to know the mathematical model of the process they control and its parameters. The ability to design different neural network architectures, such as convolutional, recurrent, or deep neural networks, makes it possible to adapt the neural regulator to the specific process it is supposed to control. On the other hand, the multiplicity of neural network architectures and their design means that we can never be sure whether a given neural network structure is optimal.
The selection of neural controller parameters is done automatically using appropriate network training algorithms.  The key element influencing the accuracy of neural regulator operation is the data used for training the neural network. The disadvantage of regulators using neural networks is the inability to demonstrate the stability of operation of the systems they control. In case of the PID regulator, despite the use of approximate models of the process, it is very often possible to prove that a closed control system will operate stably in any or a certain range of values of variables. Unfortunately, such an analysis cannot be carried out in the case of neural regulators.

In summary, the implementation of two different controllers to perform the same task provides an opportunity to learn the advantages and disadvantages of each. Students, in the course of solving the problem presented, must demonstrate theoretical knowledge of the issues and - which is particularly important for engineers - acquire the ability to practically apply theoretical knowledge.

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}